Procrastination and Dementia

### 2025-04-01 analysis/paper.qmd
Update :bug:

  <a class="btn btn-danger btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;PDF <small>(download)</small>
  </a>

  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output/manuscript+ms.pdf" target="manuscript">
  <i class="fa-solid fa-file-pdf"></i>&ensp;Manuscripty PDF
  </a>

  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript+ms.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;Manuscripty PDF <small>(download)</small>
  </a>
</p>

>>>>

  <a class="btn btn-danger btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;PDF <small>(download)</small>
  </a>
</p>

### 2025-04-01 manuscript/manuscript.qmd
Update :bug:

>>>>

# Method {#method .unnumbered}

## Data and study population {#data-and-study-population .unnumbered}

Analyses were conducted using a secondary data source; a multi-wave prospective cohort study called the Health and Retirement Study (HRS; [@Fisher2018]), which tracks the health, economic, and social well-being of over $18,000$ American adults primarily aged $>50$. The HRS is managed by the Institute for Social Research at the University of Michigan, with data collected every two years. Initial data collection of a participant is conducted through a face-to-face interview, with follow-up biennial interviews conducted either by phone or face-to-face. The average re-interview response rate ranges from $68.8\%$ to $92.3\%$ [@HRS2017]. At the time of writing, fifteen years of HRS data are currently archived.

For this study, we focused on four waves of HRS data (from 2016 to 2022). Specifically our study sample consisted of respondents who participated in an experimental module assessing procrastination during the 2020 wave. These experimental modules, administered at the end of the core HRS interview, consist of concise questionnaires designed to explore new topics or supplement existing core survey data. Each respondent receives only one experimental module, with sample sizes for each module constituting approximately 10% of the core sample. As a result, while the core HRS sample includes approximately $18,000$ respondents, our initial sample of interest consisted of $1,368$ respondents. We excluded respondents with missing cognitive assessment data for any wave $(n = 445)$ and those under 50 years of age $(n = 39)$. This resulted in a final analytic sample of $884$ respondents.

## Measures {#measures .unnumbered}

### Outcome: Cognitive Function and Cognitive Category {#outcome-cognitive-function-and-cognitive-category .unnumbered}

Cognitive function in the HRS is assessed using a series of tests adapted from the Telephone Interview for Cognitive Status (TICS; [@Brandt1988; @Fong2009]). These tests include an immediate and delayed $10$-noun free recall test, a serial seven subtraction test, and a backward count from $20$ test. Based on these assessments, @Crimmins2011 developed both a $27$-point cognitive scale and validated cut-off points to assess and classify cognitive status. Using these points, respondents who scored $12 - 27$ were classified as having normal cognition, $7 - 11$ as having MCI, and $0 - 6$ as having dementia.

### Predictor: Procrastination {#predictor-procrastination .unnumbered}

Procrastination was measured using the Pure Procrastination Scale [@steel2010], which consists of 12 items rated on a Likert scale ranging from $1$ (strongly disagree) to $5$ (strongly agree). The total procrastination score ranges from $12$ to $60$, with higher scores indicating greater procrastination. The Pure Procrastination Scale conducted in 2020 (wave 3) had a Cronbach's $\alpha$ score of $0.91$ in this sample, indicating high internal consistency. An example of a question from the scale includes "I delay making decisions until it's too late".

### Confounders {#confounders .unnumbered}

To account for potential confounding, we controlled for variables with established associations with cognitive function and procrastination. These included baseline measures conducted in 2016 (wave 1) of age, sex, educational attainment, number of cardiovascular risk factors (history of hypertension, stroke, or heart disease, and a classification on "overweight\" on a body mass index (BMI) scale $(\text{BMI} \ge 30)$), and depressive symptoms [@Abner2012; @Freedman2024; @monaghan2024a; @Yu2013]. Educational attainment was classified into three categories: no formal education, GED (General Educational Development)/high school diploma, and college/further education. Depressive symptoms were measured using an eight-item version of the Center for Epidemiological Studies Depression (CES-D) scale [@Briggs2018], with scores ranging from 0 to 8, demonstrating good internal consistency $(\alpha = 0.82)$.
